Gaussian Processes Autoregressive Models for the Disturbance Storm Time Index

### 2016-04-07 swsc.tex
Added global opt section

   \abstract{
    We aim to improve the state of the art of \emph{One Step Ahead}(OSA) prediction of the $D_{st}$ geomagnetic activity index. In the domain of OSA prediction, it is well known that $D_{st}(t)$ values correlate strongly with values one hour in the past (i.e. $D_{st}(t-1)$). While a number of forecasting models have been proposed and compared, none of them have been compared with the \emph{Persistence} model to quantify their performance improvements above this most trivial of predictors. In this study we train Gaussian Process regression models for One Step Ahead (OSE) prediction of the Disturbance Storm Time ($D_{st}$) geomagnetic index. We propose three variants of the Gaussian Process model.
    \begin{enumerate*}
      \item Gaussian Process Auto regressive GP-AR 
      \item Gaussian Process Auto regressive with eXogenous inputs GP-NARX. 
    \end{enumerate*}
    We compare the performance of these models with the current state of the art in one step ahead $D_{st}$ prediction on a set of 
    63 benchmark storms from 1998-2006. The proposed Gaussian Process models not only outperform the rest but also the \emph{Persistence} model.
    }

@@

The covariance structure of a \emph{stochastic process} is represented by a symmetric positive definite function as given in \ref{eq:kern}, by \emph{Kolmogorov's extension theorem} (see \citet{tao2011introduction}), the function $K: \mathbb{R}^d \times \mathbb{R}^d \rightarrow \mathbb{R}$ uniquely determines the process. Therefore this covariance function (also called kernel function in machine learning) helps us to specify the class of functions represented by the process.

@@

@@

@@

\subsection{Gaussian Process Nonlinear Auto-Regressive (NAR)}

@@

\subsection{Gaussian Process Nonlinear Auto-Regressive with eXogenous inputs (NARX)}

>>>>

   \abstract{
    We aim to improve the state of the art of \emph{One Step Ahead}(OSA) prediction of the $D_{st}$ geomagnetic activity index. In the domain of OSA prediction, it is well known that $D_{st}(t)$ values correlate strongly with values one hour in the past (i.e. $D_{st}(t-1)$). While a number of forecasting models have been proposed and compared, none of them have been compared with the \emph{Persistence} model to quantify their performance improvements above this most trivial of predictors. In this study we train Gaussian Process regression models for One Step Ahead (OSE) prediction of the Disturbance Storm Time ($D_{st}$) geomagnetic index. We propose three variants of the Gaussian Process model.
    \begin{enumerate*}
      \item Gaussian Process Auto regressive GP-AR 
      \item Gaussian Process Auto regressive with eXogenous inputs GP-ARX. 
    \end{enumerate*}
    We compare the performance of these models with the current state of the art in one step ahead $D_{st}$ prediction on a set of 
    63 benchmark storms from 1998-2006. The proposed Gaussian Process models not only outperform the rest but also the \emph{Persistence} model.
    }

@@

The covariance structure of a \emph{stochastic process} is represented by a symmetric positive definite function as given in \ref{eq:kern}. By \emph{Kolmogorov's extension theorem} (\citet{tao2011introduction}), the function $K: \mathbb{R}^d \times \mathbb{R}^d \rightarrow \mathbb{R}$ uniquely determines the process. Therefore this covariance function (also called kernel function in machine learning) helps us to specify the class of functions represented by the process.

@@

In this study, we construct Gaussian Process models with polynomial kernels for \emph{one step ahead} prediction of the $D_{st}$ index.

@@

With kernel based prediction models one always encounters the crucial question of what values to assign to the kernel hyper-parameters. This is known as \emph{model tuning} in kernel research parlance. There are many methods to perform \emph{model tuning}, one generally formulates an objective function and aims to optimize it in terms of the hyper-parameters. Common examples of objective functions include data likelihood, cross validation performance scores and validation set performance scores among others. Since the objective function for the model tuning problem is non-convex and non-smooth in general, gradient free approaches are used to find approximate values of local optima.

The simplest way to perform model tuning is called \emph{grid search} which entails evaluating the above objective function on a fixed grid which is constructed by making a list of values for each hyper-parameter and taking the combinations of all such values. There are more sophisticated methods for \emph{model tuning} such as \emph{coupled simulated annealing} (\citet{Xavier-De-Souza2010}) and \emph{simplex search} (\citet{Nelder1965}) but they are computationally more expensive. In this study we chose \emph{grid search} as the model tuning routine.

@@

\subsection{Gaussian Process Nonlinear Auto-Regressive (GP-AR)}

@@

\subsection{Gaussian Process Nonlinear Auto-Regressive with eXogenous inputs (GP-ARX)}
